---
output: 
  github_document:
    keep_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(thematic)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)
```

# thematic

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/thematic)](https://CRAN.R-project.org/package=thematic)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/rstudio/thematic.svg?branch=master)](https://travis-ci.org/rstudio/thematic)
[![Codecov test coverage](https://codecov.io/gh/rstudio/thematic/branch/master/graph/badge.svg)](https://codecov.io/gh/rstudio/thematic?branch=master)
<!-- badges: end -->

Simple and automatic theming of **ggplot2**, **lattice**, and **base** graphics.

## Installation

**thematic** is not yet available on [CRAN](https://CRAN.R-project.org), but you can install it now with:

``` r
remotes::install_github("cpsievert/thematic")
library(thematic)
```

## Getting started

**thematic** provides a single entry point to controling the main colors and fonts of **ggplot2**, **lattice**, and **base** graphics. To start a (global) theme, give `thematic_begin()` background/foreground/accent colors and a `font_spec()`. All of these arguments are optional, but **thematic** works best if you specify at least `bg`/`fg`/`accent`.

```{r}
library(thematic)
thematic_begin(
  bg = "#444444", fg = "#e4e4e4", accent = "#749886", 
  font = font_spec(family = "Oxanium", scale = 1.25, auto_install = TRUE)
)
```

As you'll see in the examples below, **thematic** uses these colors to inform consistent defaults not only for the plot's overall theme (e.g., plot/panel background, text color, etc), but also for `accent`ed graphical markers (e.g., **ggplot2** geom defaults), as well as `sequential` and `qualitative` color scales (these scaling defaults can also be controlled via `thematic_begin()`). `thematic_begin()` works by modifying global state in such a way that all your future plots "just work" (with [some exceptions]()), so best practice is to call `thematic_end()` when you're done using **thematic**.

## ggplot2

```{r}
library(ggplot2)
ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_raster(aes(fill = density)) + 
  geom_contour()
```

See the ggplot2 article for more examples and explanation.

### Base

Base R graphics don't have a "global" notion of `sequential` colorscales (i.e., they must be implemented as a part of each plotting call); but in most cases, you can provide the current sequential colorscale by doing something like `col = thematic_current("sequential")`:

```{r}
image(volcano)
image(volcano, col = thematic_current("sequential"))
```
 
## Lattice

```{r}
lattice::show.settings()
```


## Google fonts

There are two main controls for fonts currently: `family` and `scale`. `scale` (defaults to 1) is multiplied against all relevant font sizes. If `family` references a font that doesn't exist on the system, but the font is available on [Google Fonts](https://fonts.google.com/), **thematic** will try to download and register the font files for you (as long as `auto_install = TRUE`).

```r
font <- font_spec(family = "Oxanium", scale = 1.25, auto_install = TRUE)
thematic_begin("black", "white", font = font)
p
```

## With shiny auto-theming

**shiny** uses **thematic**'s functionality to implement it's auto-theming feature. 

TODO: an example of overriding auto-theming defaults

## Known issues

Auto-installed fonts (i.e., custom fonts downloaded from Google Fonts) currently don't work at all with the RStudio graphics device. *If* you have the **showtext** package installed, they should work with **rmarkdown**/**knitr** as well as with the default `device` in `thematic_with_device()` (the latter will also work if you have **ragg** instead of **showtext** installed):

```r
file <- thematic_with_device(plot(1), res = 144)
```

If you're in RStudio, you can preview the resulting `file` with `file.show(file)`. Moreover, to embed the `file` in **rmarkdown**/**knitr**, do `knitr::include_graphics(file)`.

### Manually installing fonts {#fonts}

If you want to use custom font(s) that aren't Google Font(s), you'll need to manually download and install/register them with R. Since **thematic** has implicit **ragg** and **showtext** integration, after downloading the font files, you can use `systemfonts::register_font()` and `sysfonts::font_add()` to register them. Another, more expensive, but permanent way to make custom fonts available to R is with `extrafont::import_font()` and `extrafont::loadfonts()`.
