---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(thematic)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)
```

# thematic

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/thematic)](https://CRAN.R-project.org/package=thematic)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

Theme **ggplot2**, **lattice**, and **base** graphics based on a few simple settings.

## Installation

**thematic** is not yet available on [CRAN](https://CRAN.R-project.org), but you can install it now with:

``` r
remotes::install_github("cpsievert/thematic")
library(thematic)
```

## Getting started

### Main colors

```{r}
library(ggplot2)
p <- ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_raster(aes(fill = density)) + 
  geom_contour()
```

```{r}
thematic_begin(bg = "darkblue", fg = "skyblue", accent = "orange")
p
```

```{r}
thematic_begin("#444444", "#e4e4e4", "#749886")
p
```

```{r}
lattice::show.settings()
```

```{r}
image(volcano)
image(volcano, col = thematic_current("sequential"))
```


```{r}
thematic_end()
p
```


### Fonts 

Coming soon?

## Caveats

1. To ensure **thematic** works properly in **knitr**/**rmarkdown**, you currently need to put `library(thematic)` (or `knitr::opts_knit$set("global.par" = TRUE)`) in a code chunk *before other code chunks* that call `thematic_begin()`.
2. **thematic** currently works best with the development version of **ggplot2**, so you may want to consider installing `remotes::install_github("tidyverse/ggplot2")`.
