---
title: "Customizing thematic"
author: "Carson Sievert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(thematic)
set.seed(101)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%", 
  fig.align = 'center',
  fig.width = 10,
  fig.asp = 0.5,
  fig.show = "hold",
  message = FALSE,
  fig.retina = 2
)
#knitr::knit_hooks$set(optipng = knitr::hook_optipng)
#knitr::opts_chunk$set(optipng = "")
```

There are two main ways to customize styling produced by **thematic**: (1) specify the global settings in `thematic_on()` and/or (2) modify defaults on the local (i.e., plot) level. 

## Global customization

There are six main components to a **thematic** theme: a background color (`bg`), a foreground color (`fg`), an `accent` color, a `font` specification, a `sequential` color palette, and a `qualitative` color palette. Any of these components can be controlled through arguments to `thematic_on()`. Here's a high level summary of the values they understand:

* `'auto'` (except for `qualitative`, which always defaults to `okabe_ito()`).
* R or CSS color codes (except for `font`, which wants a `font_spec()` object).
* `NA` (except for `bg` and `fg`), which tells **thematic** avoid setting new defaults.
* `sequential` also understands `sequential_gradient()`, which controls parameters for constructing a gradient from `bg`, `accent`, `fg` (more generally, `sequential` can be a function with `bg`, `accent`, and `fg` arguments that returns a vector of color codes).

To demonstrate how to take advantage of this behavior, consider the following plot:

```{r}
library(ggplot2)
library(patchwork)
smooth <- ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_smooth()
contour <- ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_raster(aes(fill = density)) + 
  geom_contour()
smooth + contour
```

Now, we can achieve a "dark mode" without impacting `accent` (or, by consequence, `sequential`) by setting `accent` to `NA`

```{r}
thematic_on(bg = "#222222", fg = "white", accent = NA)
last_plot()
```

And to customize `accent` without impacting the sequential color scale default:

```{r}
thematic_on(bg = "#222222", fg = "white", accent = "yellow", sequential = NA)
last_plot()
```

By default, `sequential` goes from a mixture of `fg` and `accent` to a mixture of `bg` and `accent`. 


```{r}
thematic_on(bg = "#222222", fg = "white", accent = "#6C71C4")
last_plot()
```

The amount of the mixture (i.e., weight) can be specified via `sequential_gradient()`. Here we increase the `bg` and decrease `fg` weight, effectively achieving an `accent` -> `bg` gradient: 

```{r}
thematic_on(
  bg = "#222222", fg = "white", accent = "#6C71C4", 
  sequential = sequential_gradient(bg_weight = 1, fg_weight = 0)
)
last_plot()
```

We can also flip the endpoints, which in this case gives us `bg` -> `accent` (instead of `accent` -> `bg`).

```{r}
thematic_on(
  bg = "#222222", fg = "white", accent = "#6C71C4", 
  sequential = sequential_gradient(bg_weight = 1, fg_weight = 0, fg_low = FALSE)
)
last_plot()
```

If that level of control doesn't fit your use case, can also supply a vector of color codes (or a function of `accent`, `fg`, `bg` that returns color codes):

```{r}
thematic_on(
  bg = "#222222", fg = "white", accent = "#6C71C4", 
  sequential = terrain.colors(10)
)
last_plot()
```

Similarly, `qualitative` accepts arbitrary vector of codes. In **ggplot2**, they are used so long as there are enough colors to encode the relevant number of levels (i.e., left plot). Otherwise, the default `scale_[color/fill]_hue()` is used instead (i.e., right plot).

```{r}
thematic_on(
  bg = "#222222", fg = "white", 
  qualitative = RColorBrewer::brewer.pal(8, "Dark2")
)
dark2 <- ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) + geom_point()
hue <- ggplot(mtcars, aes(wt, mpg, color = row.names(mtcars))) + geom_point()
dark2 + hue
```

It's worth noting that these examples intentionally use **ggplot2** (instead of **lattice** or **base**) to demonstrate it's relatively better handling of `qualitative` and `sequential`. As it turns out, **ggplot2** is the only approach that allows `sequential` scales to be specified independently from a particular graph's implementation, but as you'll see in the **base** graphics section below, you can still use `thematic_get_option("sequential")` to route the sequential scale to **base** (or **lattice**) graphics. Moreover, when it comes to **base** graphics `qualitative` sets the `palette()`, which gets recycled when the number of levels to encode is greater than the length of the `palette()`.

Also, note that `thematic_on()` modifies global defaults via `ggplot2::theme_set()`, `lattice::trellis.par.set()`, `graphics::par()`, and `grDevices::palette()`, which can always be modified further, or overwritten at the plot level.

## Local customization

### ggplot2 theme defaults

`thematic_on()` sets a new "complete" `theme()` default based on `bg`, `fg`, and `font` (based on a modified version of `theme_gray()`). Since this theme is "complete", you probably don't want to mix it with other complete themes (e.g., `theme_bw()`), but you can definitely override particular aspects with `theme()`.

```{r}
thematic_on(bg = "#222222", fg = "white")
p <- ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() +
  facet_wrap(~cyl)
p + theme(strip.background = element_rect(fill = "purple"))
```

By the way, it's worth noting that thematic uses a fairly arbitrary amount of mixture between the `fg` and `bg` to set the theme. If you wanted different mixture(s) of `fg` and `bg`, then `thematic_get_mixture()` is useful:

```{r}
my_theme <- theme(
  panel.background = element_rect(fill = thematic_get_mixture(0.6)),
  strip.background = element_rect(fill = thematic_get_mixture(0.3)),
  strip.text = element_text(color = thematic_get_mixture(1))
 )
p + my_theme
```

### ggplot2 geom defaults

For each relevant `Geom` (e.g., `GeomPoint`), new `Geom$default_aes` defaults are set (based on `bg`, `fg`, `accent`, and `font`). Grayscale colors (e.g., `GeomPoint$default_aes$color`) are assigned a mixture of `bg` and `fg` whereas non-grayscale colors (e.g., `GeomSmooth$default_aes$color`) are assigned the `accent` color. It's important to note these are just global defaults that only take effect if the aesthetic hasn't been specified:

```{r}
ggplot(mtcars, aes(wt, mpg)) + 
  geom_point(color = "red") +
  facet_wrap(~cyl)
```

### ggplot2 scale defaults

`thematic_on()` sets a new `scale_fill_continuous()` and `scale_color_continuous()` defaults (based on `sequential`). See the previous section for an extensive discussion on how `sequential` works and note that adding a relevant (continuous) scale renders `sequential` irrelevant:

```{r}
ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_raster(aes(fill = density)) +
  scale_fill_gradient2(midpoint = 0.02)
```

It also defines new `scale_fill_discrete()` and `scale_color_discrete()` defaults (based on `qualitative`). As with `sequential` colorscales, adding a relevant (discrete) scale renders `qualitative` irrelevant:

```{r}
ggplot(economics_long) +
  geom_line(aes(date, value01, color = variable)) +
  scale_color_brewer(type = "qual")
```

As a side note, when it comes to 3rd party **ggplot2** extension packages, **thematic** should work as expected ([let us know if it doesn't](https://github.com/rstudio/thematic)) as long as those extension packages aren't hard coding defaults in un-expected ways.

```{r, fig.asp = 0.8}
library(GGally)
ggpairs(
  iris, aes(color = Species),
  upper = list(continuous = wrap(ggally_cor, displayGrid = FALSE, alignPercent = 0.8))
)
```

### Lattice

**thematic** also works with **lattice**; however, beware that theming decisions are made so that **lattice** plots look somewhat similar to **ggplot2** (i.e. panel background is a mixture of `bg` and `fg` instead of just `bg`). Also, similar to **base** graphics, **lattice** doesn't have a global distinction between a `qualitative` and `sequential` colorscales, so `sequential` isn't used in **lattice**. Instead, for consistency with **lattice**'s default, the "regions" colorscale interpolates between `qualitative[1]`, `bg`, and `qualitative[2]`.

(Btw, for **lattice**, `accent` may be of length 2. The first is used for 'stroke' and the second for 'fill').

```{r}
# accent may be of length 2 (stroke and fill)
thematic_on(bg = "#222222", fg = "white", accent = c("purple", "red"))
library(lattice)
show.settings()
```

It might seem strange to have `bg` define the middle of the color gradient, but it's intentional so that it works well with `lattic::contourplot()` (or other cases where text wants to be placed on top of the gradient):

```{r}
library(stats)
attach(environmental)
ozo.m <- loess((ozone^(1/3)) ~ wind * temperature * radiation,
       parametric = c("radiation", "wind"), span = 1, degree = 2)
w.marginal <- seq(min(wind), max(wind), length.out = 50)
t.marginal <- seq(min(temperature), max(temperature), length.out = 50)
r.marginal <- seq(min(radiation), max(radiation), length.out = 4)
wtr.marginal <- list(wind = w.marginal, temperature = t.marginal,
        radiation = r.marginal)
grid <- expand.grid(wtr.marginal)
grid[, "fit"] <- c(predict(ozo.m, grid))
contourplot(fit ~ wind * temperature | radiation, data = grid,
            cuts = 10, region = TRUE,
            xlab = "Wind Speed (mph)",
            ylab = "Temperature (F)",
            main = "Cube Root Ozone (cube root ppb)")
detach()
```

### Base

Similar to **lattice**, **base** R graphics doesn't have a global distinction between a `qualitative` and `sequential` colorscales, it just has `palette()` (which is closest, semantically, to `qualitative`):

```{r}
par(mfrow = c(1, 2))
hist(rnorm(100))
plot(rep(1:5, each = 5), rep(1:5, 5), col = 1:25, pch = 1:25, cex = 5)
```

However, do know that you can supply the current sequential colorscale to individual plotting functions by doing something like `col = thematic_get_option("sequential")`:

```{r}
par(mfrow = c(1, 2))
image(volcano)
image(volcano, col = thematic_get_option("sequential"))
```
