---
title: "Gallery of examples"
author: "Carson Sievert"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document
---

```{r, include = FALSE}
library(thematic)
set.seed(101)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%", 
  fig.align = 'center',
  fig.width = 10,
  fig.asp = 0.5,
  fig.show = "hold",
  message = FALSE,
  fig.retina = 2
)
# Match Night Sky (the default)
library(bootstraplib)
bs_theme_base_colors(bg = "#13222E", fg = "#ACBECC")
bs_theme_accent_colors(primary = "#9770B2")
bs_theme_fonts("Fira Sans Condensed")
bs_theme_add(
  rules = "@import url('https://fonts.googleapis.com/css?family=Fira+Sans+Condensed')"
)
```

## {.tabset .tabset-pills}

Below are a few themes from <https://themer.dev/>

TODO:
 * Show the ports instead (or maybe others)?
 * Support the light variant?
 * Tab callback to change the page's styling?
 * Integrate with rest of site (ditch pkgdown?)


### Night Sky {.tabset .tabset-pills}

```{r}
thematic_on(
  bg = "#13222E", fg = "#ACBECC", accent = "#9770B2", 
  font = font_spec("Fira Sans Condensed", scale = 1.2)
)
```

#### ggplot

```{r}
library(ggplot2)
qplot(1:10)
```

#### lattice

```{r}
lattice::show.settings()
```

#### base

```{r}
image(volcano, col = thematic_get_option("sequential"))
```

### Polar Ice {.tabset .tabset-pills}

```{r}
thematic_on(
  bg = "#44484F", fg = "#C7F3FF", accent = "#94CEF2", 
  font = font_spec("Fira Sans Condensed", scale = 1.2)
)
```

```{r, child = "_gallery-examples.Rmd"}
```


### Finger Paint {.tabset .tabset-pills}

```{r}
thematic_on(
  bg = "#2A2021", fg = "#EDEBEE", accent = "#6484B5", 
  font = font_spec("Fira Sans Condensed", scale = 1.2)
)
```

```{r, child = "_gallery-examples.Rmd"}
```


### Monkey {.tabset .tabset-pills}

```{r}
thematic_on(
  bg = "#48463E", fg = "#E0E2DD", accent = "#7EB9D0", 
  font = font_spec("Fira Sans Condensed", scale = 1.2)
)
```

```{r, child = "_gallery-examples.Rmd"}
```


### Right in the Teals {.tabset .tabset-pills}

```{r}
thematic_on(
  bg = "#171D1D", fg = "#CDDEDE", accent = "#00ACBD", 
  font = font_spec("Fira Sans Condensed", scale = 1.2)
)
```

```{r, child = "_gallery-examples.Rmd"}
```
